\LARGE{{\bf{Gravitational Wave Signals from Multiple Hidden Sectors} \\ }}

### 2020-02-12 gwnnDraft02.tex
typos corrected, appendix expanded

%Explicitly, all exotic sectors (and heavy standard sectors) have strong first order phase transitions that lead to runaway bubble walls ($\alpha >\alpha_{\infty}$). 
Each component of the spectral energy density in Eq.~(\ref{eqn:SpectralEng}) is proportional to a power of their respective efficiency factors $\kappa$. The relative strength of the efficiency factors is dependent on the ratio $\frac{\alpha_{\infty}}{\alpha}$ --- since both $\alpha_{\infty}$ and $\alpha$ are parameterically small \cite{Bai:2018dxf,Helmboldt:2019pan} a range of possible scenarios can occur. Here, we discuss the two ends of this spectrum: pure runaway walls and pure non-runaway walls. The former scenario with runaway bublle walls leads to the efficiency factors for the sound wave and MHD contributions being small and ensures GWs are dominantly produced from bubble collisions, $h^2\Omega_{\textrm{GW}} \approx h^2\Omega_{\phi}$: this is what we assume for the remainder of this section. In the latter case, bubbles are non-runaway (but $v_w \sim 1$ still \cite{Breitbach:2018ddu,Bodeker2017}) such that non-bubble collision contributions being important --- ultimately leading to significant changes to the GW profile. This case and the gravitational waves it produces are examined in the appendix. Intermediate results are of course possible, and would feature profiles somewhere in between the two extremes.

@@

In the former case, this demonstrates the power of GW signals to probe deep into the unknown arena of complex hidden sectors. Individual SFOPT are understood to create GW that are assumed to follow an approximate power law. If a model predicts of the presence of 2, 5, or more additional sectors, or features a single extra sector with multiple PTs, deviations from a standard power law can occur. The multiple transitions that occur in the models outlined here create signals that follow this trend: although the individual GW do obey approximate power laws, their sum does not --- leading to a unique signal indicating so-called dark complexity. Explicitly, a broadening or distortion of the signal around the peak frequency, precisely where the signal has the most energy, could point to a multi-SFOPT scenario and gently guide us in the direction of multiple hidden sectors. %ultimately providing another means to test and constrain theories.

@@

\begin{figure*}[tb]
\centering
\begin{minipage}[c]{\textwidth}
\includegraphics[width=.45\textwidth ]{MaxSW.png}
\hfill
\includegraphics[width=.45\textwidth]{largeSW.png} 
\hfill
\includegraphics[width=.45\textwidth]{MedSW.png} 
\hfill
\includegraphics[width=.45\textwidth]{fiveSW.png} 
\end{minipage}
\hfill
\caption{ Gravitational wave spectral energy density for the various scenarios found in Tab.~(\ref{tab:cases}).  All contributions are assumed to be from a non-runaway phase transition with terminal velocity $v = 0.95$. The colored solid lines use $\beta/H  = 10$ where as the dashed gray line is the total contribution of all sectors for $\beta/H  = 10^4$ . The top left figure shows the individua sound wave and MHD contributions.  The shaded curves are the same as Fig.~\ref{fig::Nnatural}. In contrast to the runaway case, most scenarios evade the projected sensistivies of future experiments and therefore is the less optimal situation }
\label{fig:Haasw}
\end{figure*}

@@

\begin{equation}
\begin{split}
h^2\Omega_{v}^* = 1.6 \times 10^{-1} \, v  \, \bigg( \frac{\kappa_{v}\;  \alpha}{1 + \alpha} \bigg)^2 \bigg( \frac{H}{\beta} \bigg)^1  S_{v}(f),
\\
h^2\Omega_{turb}^* = 2.01\times 10^{1} \, v \, \bigg( \frac{\kappa_{turb} \; \alpha}{1 + \alpha} \bigg)^{3/2} \bigg( \frac{H}{\beta} \bigg)^1  S_{turb}(f).
\end{split}
\end{equation}
Where unlike the runaway case, we do not assume $v = 1$ due to the bubbles reaching a terminal velocity. The MHD effieceny factor is a fraction of the sound waves, $\kappa_{turb} = \epsilon\kappa_{v}$. Current similations have motivated a range of $\epsilon \sim 0.05 - 0.10$  \citep{Caprini:2015zlo}, where we take the optimistic case of $\epsilon = 0.10$. Both contributions have unique spectral shapes, given to be \citep{Breitbach:2018ddu},

\begin{equation}
\begin{split}
S_{v}(f) = (f/f_{p,v})^3\bigg(\frac{7}{4+3(f/f_{p,v})^2}\bigg)^{7/2},
\\
S_{turb}(f) = \frac{(f/f_{p,turb})^{3}}{(1 + f/f_{p,turb})^{11/3} (1 + 8\pi(f/H))}.
\end{split}
\end{equation}
The MHD energy density has a spectral shape dependent on the hubble rate at the time of nucleation, $H$.  Similar to the scalar spectral shape in Eq.~(\ref{eqn::spectralshape}), the frequencies are scaled by their respective temperature dependent peak frequency,

@@

We evolve the frequencies and energy densities with the same redshift factors, Eq.~(\ref{eqn::redshifting}) from Sec~\ref{sec:gw}. Fig.~\ref{fig:Haasw} shows the sum of the GW energy density for the scenarios in Table~\ref{tab:cases}, but instead for the non-runaway case, with a terminal velocity of $v = 0.95$.

>>>>

%Explicitly, all exotic sectors (and heavy standard sectors) have strong first order phase transitions that lead to runaway bubble walls ($\alpha >\alpha_{\infty}$). 
Each component of the spectral energy density in Eq.~(\ref{eqn:SpectralEng}) is proportional to a power of their respective efficiency factors $\kappa$. The relative strength of the efficiency factors is dependent on the ratio $\frac{\alpha_{\infty}}{\alpha}$ --- since both $\alpha_{\infty}$ and $\alpha$ are parameterically small \cite{Bai:2018dxf,Helmboldt:2019pan}, a range of possible scenarios can occur. Here, we discuss the two ends of this spectrum: pure runaway walls and pure non-runaway walls. The former scenario with runaway bublle walls leads to the efficiency factors for the sound wave and MHD contributions being small and ensures GWs are dominantly produced from bubble collisions, $h^2\Omega_{\textrm{GW}} \approx h^2\Omega_{\phi}$: this is what we assume for the remainder of this section. In the latter case, bubbles are non-runaway (but $v_w \sim 1$ still \cite{Breitbach:2018ddu,Bodeker2017}) such that non-bubble collision contributions being important --- ultimately leading to significant changes to the GW profile. This case and the gravitational waves it produces are examined in the appendix. Intermediate results are of course possible, and would feature profiles somewhere in between the two extremes.

@@

In the former case, this demonstrates the power of GW signals to probe deep into the unknown arena of complex hidden sectors. Individual SFOPT are understood to create GW that are assumed to follow an approximate power law. If a model predicts the presence of 2, 5, or more additional sectors, or features a single extra sector with multiple PTs, deviations from a standard power law can occur. The multiple transitions that occur in the models outlined here create signals that follow this trend: although the individual GW do obey approximate power laws, their sum does not --- leading to a unique signal indicating so-called dark complexity. Explicitly, a broadening or distortion of the signal around the peak frequency, precisely where the signal has the most energy, could point to a multi-SFOPT scenario and gently guide us in the direction of multiple hidden sectors. %ultimately providing another means to test and constrain theories.

@@

\begin{figure*}[tb]
\centering
\begin{minipage}[c]{\textwidth}
\includegraphics[width=.45\textwidth ]{MaxSW.png}
\hfill
\includegraphics[width=.45\textwidth]{largeSW.png} 
\hfill
\includegraphics[width=.45\textwidth]{MedSW.png} 
\hfill
\includegraphics[width=.45\textwidth]{fiveSW.png} 
\end{minipage}
\hfill
\caption{ Gravitational wave spectral energy density for the various scenarios found in Tab.~(\ref{tab:cases}).  All contributions are assumed to be from a non-runaway phase transition with terminal velocity $v = 0.95$. The coloured solid lines use $\beta/H  = 10$ where as the dashed grey line is the total contribution of all sectors for $\beta/H  = 10^4$ . The top left figure shows the individual sound wave and MHD contributions.  The shaded curves are the same as Fig.~\ref{fig::Nnatural}. In contrast to the runaway case, most scenarios evade the projected sensitivities. }
\label{fig:Haasw}
\end{figure*}

@@

\begin{equation}
\begin{split}
h^2\Omega_{v}^* = 1.6 \times 10^{-1} \, v  \, \bigg( \frac{\kappa_{v}\;  \alpha}{1 + \alpha} \bigg)^2 \bigg( \frac{H}{\beta} \bigg)^1  S_{v}(f),
\\
h^2\Omega_{turb}^* = 2.01\times 10^{1} \, v \, \bigg( \frac{\kappa_{turb} \; \alpha}{1 + \alpha} \bigg)^{3/2} \bigg( \frac{H}{\beta} \bigg)^1  S_{turb}(f).
\end{split}
\end{equation}
Unlike the runaway case, we do not assume $v = 1$ due to the bubbles reaching a terminal velocity. The MHD efficiency factor is a fraction of the sound waves, $\kappa_{turb} = \epsilon\kappa_{v}$. Current simulations have motivated a range of $\epsilon \sim 0.05 - 0.10$  \citep{Caprini:2015zlo}, where we take the optimistic case of $\epsilon = 0.10$. Both contributions have unique spectral shapes, given to be \citep{Breitbach:2018ddu}, 
%
\begin{equation}
\begin{split}
S_{v}(f) = (f/f_{p,v})^3\bigg(\frac{7}{4+3(f/f_{p,v})^2}\bigg)^{7/2},
\\
S_{turb}(f) = \frac{(f/f_{p,turb})^{3}}{(1 + f/f_{p,turb})^{11/3} (1 + 8\pi(f/H))}.
\end{split}
\end{equation}
The MHD energy density has a spectral shape dependent on the Hubble rate at the time of nucleation, $H$.  Similar to the scalar spectral shape in Eq.~(\ref{eqn::spectralshape}), the frequencies are scaled by their respective temperature dependent peak frequency,

@@

We evolve the frequencies and energy densities with the same redshift factors, Eq.~(\ref{eqn::redshifting}) from Sec~\ref{sec:gw}. Fig.~\ref{fig:Haasw} shows the sum of the GW energy density for the scenarios in Table~\ref{tab:cases}, but instead for the non-runaway case, with a terminal velocity of $v = 0.95$. For this case we see that the spectra tend to be shifted to lower frequencies and are more likely to fall in the gap between the interferometers and the pulsar based detectors. On the other hand, larger values of $\beta/H$ increase the typical frequency, so this case becomes more sensitive in some scenarios to values of $\beta/H$ on the larger end of the considered range.

Finally, we note that because the sound wave and MHD contributions have different spectral shapes, the overall spectrum has a kink at a frequency above the peak. In the top left panel of Fig.~\ref{fig:Haasw}, we show the two contributions separately in addition to their sum to highlight this effect.
